---
title: "Assignment 6"
author: "Rex Shen"
date: "10/30/2021"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, messages = FALSE)
```


### Part 1

```{r, warning = FALSE, messages = FALSE, echo = FALSE, include = FALSE}
# Load Packages
library(mapview)
library(censusapi)
library(tidyverse)
library(tigris)
library(sf)
library(leaflet)
```


```{r, warning = FALSE, messages = FALSE, echo = FALSE, include = FALSE}
Sys.setenv(CENSUS_KEY="c8aa67e4086b4b5ce3a8717f59faa9a28f611dab")

Pums_2019_1Yr = getCensus(
   name = "acs/acs1/pums",
   vintage = 2019,
   region = "public use microdata area:*",
   regionin = "state:06",
   vars = c(
     "SERIALNO",
     "SPORDER",
     "PWGTP",
     "WGTP",
     "YBL",
     "BLD",
     "TEN",
     "MV",
     "HINCP",
     "AGEP"
   )
 )

CA_Pumas =
  pumas("CA", cb = T, progress_bar = F)

SF_Boundary =
  counties("CA", cb = T, progress_bar = F) %>%
  filter(NAME == "San Francisco")

SF_Pumas =
  CA_Pumas %>% 
  st_centroid() %>% 
  .[SF_Boundary, ] %>% 
  st_drop_geometry() %>% 
  left_join(CA_Pumas %>% select(GEOID10)) %>% 
  st_as_sf()
```

```{r}
mapview(SF_Pumas)
```



```{r}
SF_Pums =
  Pums_2019_1Yr %>% 
  mutate(
    PUMA = str_pad(public_use_microdata_area, 5, "left", "0")
  ) %>% 
  filter(PUMA %in% SF_Pumas$PUMACE10)


SF_Pums_Clean = SF_Pums %>% mutate(YBL = as.numeric(YBL), AGEP = as.numeric(AGEP), HINCP = as.numeric(HINCP)) %>%
  filter(YBL %in% 1:3) %>%
  arrange(AGEP) %>%
  group_by(SERIALNO) %>% summarize_all(first) %>% mutate(Leadrisk = ifelse(HINCP < 90000 &
  AGEP < 6, 1, 0))
```


```{r}
SF_Pums_Clean = SF_Pums_Clean %>% mutate(
    BLD = BLD %>% 
      factor(
        levels = SF_Pums_Clean$BLD %>% 
          unique() %>%
          as.numeric() %>% 
          sort()
      )) %>% mutate(
    TEN = TEN %>% 
      factor(
        levels = SF_Pums_Clean$TEN %>% 
          unique() %>%
          as.numeric() %>% 
          sort()
      )) %>% mutate(MV = MV %>% 
      factor(
        levels = SF_Pums_Clean$MV %>% 
          unique() %>%
          as.numeric() %>% 
          sort()
      ))

SF_Pums_Clean$PUMA = factor(SF_Pums_Clean$PUMA)
```


```{r}
Logit_Model = glm(
  Leadrisk ~ BLD + TEN + MV + PUMA,
  family = quasibinomial(),
  data = SF_Pums_Clean
)

summary(Logit_Model)
```

```{r}
set.seed(1)
RandomRow = data.frame(sample_n(SF_Pums_Clean, size = 1))
print(RandomRow)
print(as.numeric(predict(Logit_Model, RandomRow, type = "response")))
```


From our random record, we see that the predictor variables BLD, TEN, MV, and PUMA are $3$, $3$, $4$, and $07506$ respectively. Finally, the predicted score is $\approx 0.0815$.

### Part 2

```{r}
Scores = as.vector(predict(Logit_Model, data.frame(SF_Pums_Clean), type = "response"))

Indices = c()
for (i in 1:length(Scores)) {
  if (Scores[i] >= 0.10) {
    Indices = append(Indices, i)
  }
}

Subset = SF_Pums_Clean[Indices, ]
Subset = Subset[, c("BLD", "TEN", "MV")]
Summary_2x2 =
  SF_Pums_Clean %>% 
  mutate(
    Leadrisk = ifelse(
      Leadrisk == 1, 
      "Risk", 
      "No Risk"
    )
  ) %>% 
  pull(Leadrisk) %>% 
  table(Scores >= 0.10)

Summary_2x2
```

### Analysis/Interpretation

For the scores that exceeded the threshold, the predictor variable BLD was mostly 2, indicating that the units in structure were one family house detached. TEN was mostly 3 and 4, indicating that the housing was rented and occupied without payment or rent respectively. MV was mostly 3 and 4, indicating that when moved into the house or apartment, most of the individuals stayed in for $2-4$ and $5-9$ years respectively.

If we apply this model in real-life, we would expect $27 + 4 = 31$ mail outs based on threshold score, and we expect $\frac{4}{31} * 100 \approx 12.9\%$ to actually go to low income households with children. Moreover, we would expect $\frac{4}{33 + 4} * 100 \approx 10.8\%$ of all at-risk children to be reached.

In terms of ways to improve strategy assuming we don't directly have access to information about incomes and presence of children at the address level, we could use a proxy, such as type of occupation or total number of people (excluding children) in a household. We could also improve our strategy by applying a different model based on what our response is, such as logistic regression.

