---
title: "Assignment 4"
author: "Rex Shen"
date: "10/15/2021"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include = FALSE}
library(knitr)
knitr::opts_chunk$set(echo = F, message = FALSE, warning = FALSE)
opts_chunk$set(tidy.opts = list(width.cutoff = 40), tidy = TRUE)
```


### Part 1


```{r}
library(tidyverse)
library(censusapi)
library(base)

Sys.setenv(CENSUS_KEY="c8aa67e4086b4b5ce3a8717f59faa9a28f611dab")

acs_vars_2019_5yr <-
  listCensusMetadata(
    name = "2019/acs/acs5",
    type = "variables"
  )
```


```{r}
Total_Education_Attainment = NULL

Race = c("White Alone", "Black or African American Alone", "American Indian and Alaska Native Alone", "Asian Alone",  "Native Hawaiian and Other Pacific Islander Alone", "Some Other Race Alone", "Two Or More Races")


for (i in 1:length(Race)) {
  
  Education_Attainment <-
  getCensus(
    name = "acs/acs5",
    vintage = 2019,
    region = "county:013", # Contra Costa County 
    regionin = "state:06",
    vars = paste0("group(C15002", LETTERS[i],")")
  ) %>%
  select(!c(GEO_ID,state,NAME) & !ends_with(c("EA","MA","M"))) %>%
  pivot_longer(
    ends_with("E"),
    names_to = "name",
    values_to = "estimate"
  ) %>%
  left_join(
    acs_vars_2019_5yr %>% 
      select(name, label)
  ) %>% 
  select(-name) %>% 
  separate(
    label,
    into = c(NA, NA, NA, "Education_Attainment"),
    sep = "!!"
  ) %>% mutate(race = Race[i])
  
  Education_Attainment <- Education_Attainment[-c(2, 7), ]
  Education_Attainment[1, 3] = "Total"
  
  Total_Education_Attainment = rbind(Total_Education_Attainment, Education_Attainment)
}
```


```{r}
Total_Education_Attainment = Total_Education_Attainment %>% 
  group_by(Education_Attainment, race) %>% 
  summarize(estimate = sum(estimate))

Counter = 0
for (i in 1:28) {
  
  if (29 + Counter > nrow(Total_Education_Attainment)) {
    Counter = 0
  }
    Total_Education_Attainment$Proportion[i] =
      Total_Education_Attainment$estimate[i]/Total_Education_Attainment$estimate[29 + Counter]
    Counter = Counter + 1
}

for (i in 29:nrow(Total_Education_Attainment)) {
  Total_Education_Attainment$Proportion[i] =
    Total_Education_Attainment$estimate[i]/sum(Total_Education_Attainment$estimate[29:35])
}
```



```{r}
newPlot = ggplot(data = Total_Education_Attainment) +
  geom_bar(
    aes(
      x = Education_Attainment,
      y = Proportion,
      fill = race
    ),
    stat = "identity",
    position = "fill"
  ) +
  labs(
    x = "Educational Attainment",
    y = "Proportion of Population 25 Yrs or Older",
    title = "Contra Costa County Educational Attainment",
    fill = "Race"
  ) + coord_flip() + theme(
    legend.position = "bottom",
    legend.direction = "vertical"
  )  +
  guides(
    fill = guide_legend(
      reverse = T
    )
  )

newPlot
```


### Analysis and Observations

Note that we are looking at Contra Costa County exclusively.


In the plot, we see that "White Alone" takes up a significant proportion of the total numbers. It appears that there is a relative difference in spread among the different races for most of the categories. For instance, in the Bachelor's Degree or Higher, there is more proportion in the "Asian Alone" and "White Alone" than other races, where "Asian Alone" and "White Alone" take roughly the same proportion. Moreover, for Less Than High School Diploma, there is a higher proportion of Some Other Race Alone than other categories. On the other hand, there appears to be a somewhat even spread for Some College or Associate's Degree, whereas there is less even spread in the others. Overall, there is disproportionate education attainment in Contra Costa County in most of the categories of educational attainment for those 25+.

### Part 2

In this section, we obtain three categories: "Latinx", "White, Not Latinx", and "Not White, Not Latinx." That is, we perform subtraction to obtain the proportion of "Not White, Not Latinx" using the 2019 1-Yr Data for Contra Costa County for those 25+.

```{r}
# Get Hispanic or Latino and White Alone, Not Hispanic or Latino and White Alone

Race = c("Hispanic or Latino", "White Alone, Not Hispanic or Latino")
Letters = c("I", "H")
  
Education_Attainment <-
  getCensus(
    name = "acs/acs1",
    vintage = 2019,
    region = "county:013",
    regionin = "state:06",
    vars = c("C15002A_001E", "C15002A_002E", "C15002A_003E", "C15002A_004E", "C15002A_005E",
             "C15002A_006E", "C15002A_007E", "C15002A_008E", "C15002A_009E", "C15002A_010E",
             "C15002A_011E", "C15002I_001E", "C15002I_002E", "C15002I_003E", "C15002I_004E", "C15002I_005E","C15002I_006E", "C15002I_007E", "C15002I_008E", "C15002I_009E", "C15002I_010E",
             "C15002I_011E", "C15002H_001E", "C15002H_002E", "C15002H_003E", "C15002H_004E", "C15002H_005E", "C15002H_006E", "C15002H_007E", "C15002H_008E", "C15002H_009E", "C15002H_010E",
             "C15002H_011E") 
  ) %>% select(!c(state)) %>%
pivot_longer(
    ends_with("E"),
    names_to = "name",
    values_to = "estimate"
  ) %>%
  left_join(
    acs_vars_2019_5yr %>% 
      select(name, label)
  )

for (i in 1:nrow(Education_Attainment)) {
  
  if (grepl("H", Education_Attainment$name[i])) {
    Education_Attainment$race[i] = "White Alone, Not Hispanic or Latino"
  } else if (grepl("I", Education_Attainment$name[i])) {
    Education_Attainment$race[i] = "Hispanic or Latino"
  } else {
    Education_Attainment$race[i] = "White Alone"
  }
}

  
  Education_Attainment <- Education_Attainment[-c(2, 7, 13, 18, 24, 29), ]
  
  Education_Attainment = Education_Attainment %>% separate(
    label,
    into = c(NA, NA, NA, "Education_Attainment"),
    sep = "!!"
  )
  
  Education_Attainment = Education_Attainment[, -c(2)]
  
  Education_Attainment[1, 3] = "Total"
  Education_Attainment[10, 3] = "Total"
  Education_Attainment[19, 3] = "Total"
  
  Education_Attainment = Education_Attainment %>% group_by(Education_Attainment, race) %>%
    summarize(estimate = sum(estimate))
```

```{r}

Education_Attainment$Proportion = NA

Counter = 0
for (i in 1:12) {
  
  if (13 + Counter > nrow(Education_Attainment)) {
    Counter = 0
  }
  if (Education_Attainment$race[i] != "White Alone") {
    
    Education_Attainment$Proportion[i] =
      Education_Attainment$estimate[i]/Education_Attainment$estimate[13 + Counter]
    Counter = Counter + 1
  }
    
}


for (i in 1:nrow(Education_Attainment)) {
  
  if (Education_Attainment$race[i] == "White Alone") {
    Education_Attainment$Proportion[i] = 1 - Education_Attainment$Proportion[i - 1] - 
      Education_Attainment$Proportion[i + 1]
    
    Education_Attainment$race[i] = "Not White, Not Latinx"
  }
  
}

Education_Attainment = Education_Attainment[-c(13, 14, 15), ]
```



```{r}
# Plot Data
newPlot = ggplot(data = Education_Attainment) +
  geom_bar(
    aes(
      x = Education_Attainment,
      y = Proportion,
      fill = race
    ),
    stat = "identity",
    position = "fill"
  ) +
  labs(
    x = "Educational Attainment",
    y = "Proportion of Population 25 Yrs or Older",
    title = "Contra Costa County Educational Attainment",
    fill = "Ethnicity"
  ) + coord_flip() + theme(
    legend.position = "bottom",
    legend.direction = "vertical"
  )  +
  guides(
    fill = guide_legend(
      reverse = T
    )
  )

newPlot

```

### Analysis and Observations

It appears there is a more apparent disproportionate representation of different sub-populations than before. Particularly, "Hispanic or Latino" consistently has a smaller proportion in the plot. On the other hand, "Not White, Not Latinx" has a greater proportion in the plot. It appears that the spread is more even in the category of "Some College or Associate's Degree," but less so in the other categories. Perhaps, this may suggest that there are some confounding factors at play.

